\documentclass[12pt, letterpaper, preprint]{aastex} 

\usepackage{graphicx}
\usepackage{amsmath}

\newcommand{\tmin}[1][]{t_{\mathrm{min} #1}}
\newcommand{\tmax}[1][]{t_{\mathrm{max} #1}}
\newcommand{\amin}{a_{\mathrm{min}}}
\newcommand{\amax} {a_{\mathrm{max}}}

\newcommand{\dlt}{\Delta\log t_i}
\newcommand{\dt}{\Delta t_i}

\newcommand{\tintegral}{\int_{\tmin[,i]}^{\tmax[,i]} dt}
\newcommand{\tinterval}{\right|_{\tmin[,i]}^{\tmax[,i]}}
\newcommand{\clip}[3][]{{#1}_{\top {#2}}^{\bot {#3}}}

\newcommand{\sftrunc}{t_{\mathrm{trunc}}}
\newcommand{\tage}{t_{\mathrm{age}}} 
\begin{document}
\author{B. Johnson}

%\begin{center}
%\today
%\end{center}

\section{Introduction}
We wish to find the proper weighting of discrete SSPs to reproduce the spectrum $f$ from SFR occurring within some bin of  lookback times $(\amin, \amax)$.
We approximate the spectrum $f(t)$ at any lookback time $t$ as a linear interpolation of the (total stellar mass normalized) spectra at the discrete SSP lookback times $t_i$, for $i=1,...,N$ 
\begin{eqnarray}
f & = & \int_{\amin}^{\amax} dt \, f(t) \, SFR(t) \nonumber \\
f(t) & = & w_{i}(t) \, f_i + w_{i+1}(t) \, f_{i+1} \, , \quad t_i < t < t_{i+1}  \nonumber \\
w_{i}(t) & = & \frac{t_{i+1} - t}{t_{i+1}  - t_i} \nonumber \\
w_{i+1}(t) & = & 1 - w_i(t) \quad . \nonumber
\end{eqnarray}
Then we can write the \emph{total} contribution of SSP $i$ to the total spectrum as a sum of the integrated weight due to SF in the sub-bin to the ``right'' (higher lookback times) and the  sub-bin to the ``left'' (lower lookback times).
The sub-bins are defined by the spacing of the SSP $t_i$ values. 
The limits of each integral are determined either by the spacing of the SSP points $t_i$ or, when the bin does not completely span the sub-bin, by the $\amin$ and/or $\amax$ values.
\begin{eqnarray}
W_i & = &  \tintegral \, \mathrm{SFR}(t) \, w_i(t) \nonumber \\
 & = & W_{i, \mathrm{right}} + W_{i, \mathrm{left}} \nonumber \\
W_{i, \mathrm{right}} & = & \tintegral \, \mathrm{SFR}(t) \, \frac{t_{i+1} - t}{t_{i+1}  - t_i} \nonumber \\
W_{i, \mathrm{left}} & = & \int_{\tmin[,i-1]}^{\tmax[,i-1]} dt \, \mathrm{SFR}(t) \, \frac{t - t_{i-1}}{t_{i}  - t_{i-1}} \nonumber \\
\tmin[,i] & = & \clip[\amin]{t_i}{t_{i+1}} \nonumber \\
\tmax[,i] & = & \clip[\amax]{t_i}{t_{i+1}} \nonumber
\end{eqnarray}
where $\clip[x]{a}{b}$ denotes $x$ \emph{clipped} to the interval $(a,b)$. If we take $\mathrm{SFR}(t) = 1$ then we can define
\begin{eqnarray}
g_i & \equiv & \tintegral \, \mathrm{SFR}(t) \, \frac{t_{i+1} - t}{t_{i+1}  - t_i} \nonumber  \\
 & = & \left. \frac{1}{t_{i+1} - t_i} \, \left[ t_{i+1} \, t - \frac{t^2}{2} \right] \tinterval
\end{eqnarray}
which leads to
\begin{eqnarray}
W_i & = & g_{i} - g_{i-1}
\end{eqnarray}
and one only has to calculate the ${g_i}$ once and then take differences between shifted versions to obtain the total weights.

This formalism works even if $\amin$ and $\amax$ both fall between the same $t_i$ and $t_{i+1}$; that is, if the desired bin is narrower than the SSP defined sub-bins.
It also works if a sub-bin falls completely outside the $(amin, amax)$ interval.
This is because the clipping will cause $\tmin$ and $\tmax$ to be the same for bins not at least partially within the $(amin, amax)$ interval, and thus all integrals will evaluate to zero.

\section{Interpolation in $\log t$}

For the case of a linear interpolation in $\log t$, which is probably more appropriate, we can write
\begin{eqnarray}
w_{i}(t) & = & \frac{ \log t_{i+1} - \log t}{ \log t_{i+1}  - \log t_i} \nonumber
\end{eqnarray}
in the first set of equations above. The we simply replace $g_i$ by $s_i$ defined as 
\begin{eqnarray}
s_i & \equiv & \tintegral \, \mathrm{SFR}(t) \, \frac{\log t_{i+1} - \log t}{\log t_{i+1}  - \log t_i} \nonumber \\
 & = & \left. \frac{1}{\log t_{i+1} - \log t_i} \left[ t \, \log t_{i+1} - t \log\left(\frac{t}{e}\right) \right] \tinterval
\end{eqnarray}
to obtain a total weight for SSP $i$ of
\begin{eqnarray}
W_i & = & s_{i} - s_{i-1} \nonumber
\end{eqnarray}

The key in both cases is to properly calculate $\tmax$ and keep track of the indices 
(especially for the first and last SSP, where $s_{i-1}$ and $s_i$ are not defined, respectively.)  

\section{Lookback Time Zero}

\section{More complex SFHs}
For more complicated SFH the integrals in the basic definition of $s_i$ must be calculated again by hand, (remembering here that $t$ is \emph{lookback time}.)  
We define $s_i$ for interpolation in $\log t$ and $g_i$ for interpolation in $t$.
\begin{eqnarray}
s_i & \equiv & \tintegral \, \mathrm{SFR}(t) \, \frac{\log t_{i+1} - \log t}{\dlt} \\
g_i & \equiv & \tintegral \, \mathrm{SFR}(t) \, \frac{t_{i+1} - t}{\dt}
\end{eqnarray}
where for brevity we've defined in this section
\begin{eqnarray}
\dt & \equiv & t_{i+1} -  t_i  \nonumber \\
\dlt & \equiv & \log t_{i+1} - \log t_i \nonumber
\end{eqnarray}
In general the limits of the integral ($\tmin, \tmax$) will be defined by the SSP temporal grid $\{t_i\}$ as well as any discontinuities in the derivative of the SFR (e.g. the edges of step function already enocountered).

For convenience we will also note that the integral
\begin{eqnarray}
H_i & \equiv & \tintegral \, e^{t/\tau} \, \log t \nonumber \\
  & = & \left. \tau \left[ \log t \, e^{t/\tau} - \log(e) \, \mathrm{Ei}(t/\tau) \right]\tinterval
\end{eqnarray}
where Ei is the exponential integral.

\subsection{$\tau$-model}
Here we have a SFH defined by three parameters
\[ 
\mathrm{SFR}(t) = 
\begin{cases}
K \, e^{-T/\tau}, &  \text{if } 0 < T \leq t_{\mathrm{trunc}} \\
0, & \text{otherwise}
\end{cases}
\]
\begin{eqnarray}
T & = & \tage - t \nonumber \\
t_q & = & \tage - t_{\mathrm{trunc}} \nonumber 
\end{eqnarray}
where $T$ is defined as the time since the start of star formation, \emph{not} lookback time.  
The lookback time  (in the reference frame of the galaxy) is given by $t$, which goes from 0 to $\tage$.
The parameters are: 
$\tage$, the age of the galaxy since the start of SF (\texttt{tage} in FSPS), 
$\tau$, the $e$-folding time (\texttt{tau} in FSPS), and
$\sftrunc$, the time since the start of SF that truncation occurs (\texttt{sf\_trunc} in FSPS).
Note that in the current FSPS definition, $t_{\mathrm{trunc}}=0$ corresponds to no truncation, so it is actually $t_{\mathrm{trunc}} = \tage$ and $t_q=0$.
The FSPS \texttt{sf\_start} is basically redundant with \texttt{tage} and confusing so we ignore it.

In this case we have, for the logarithmic time interpolation,
\begin{eqnarray}
s_i  & = & \tintegral \, K \, e^{(t-\tage) /\tau}\, \frac{\log t_{i+1} - \log t}{\dlt} \nonumber \\
      & = & \frac{K \, e^{-\tage/\tau}}{\dlt} \left[ \left. \tau \, \log t_{i+1} \, e^{t/\tau} \tinterval -  H_i\right] \nonumber \\
\tmin[,i] & = & \clip[t_q]{t_i}{t_{i+1}} \nonumber \\
\tmax[,i] & = & \clip[\tage]{t_i}{t_{i+1}} \quad . \nonumber
\end{eqnarray}
Substituting for $H_i$ we have, finally, 
\begin{eqnarray}
s_i & = & \left. \frac{K \, \tau \, e^{-\tage/\tau}}{\dlt} \left [(\log t_{i+1} - \log t)\, e^{t/\tau} + 
          \log(e) \, \mathrm{Ei}(t/\tau)\right]  \tinterval
\end{eqnarray}
For interpolation in linear time the integral is a bit more straightforward
\begin{eqnarray}
g_i & = & \left. \frac{K \, \tau \, e^{-\tage/\tau}}{\dt} (t_{i+1} - t + \tau)\, e^{t/\tau} \tinterval
\end{eqnarray}
where the limits $\tmin$ and $\tmax$ are defined as above.
One can either calculate the normalization constant $K$ analytically, or simply renormalize the final weights $W_i$ by their sum.

\subsection{Delayed-$\tau$}
Here the SFH is also defined by three parameters
\[ 
\mathrm{SFR}(t) = 
\begin{cases}
K \, T \, e^{-T/\tau}, &  \text{if } 0 < T \leq t_{\mathrm{trunc}} \\
0, & \text{otherwise}
\end{cases}
\]
\begin{eqnarray}
T & = & \tage - t \nonumber \\
t_q & = & \tage - t_{\mathrm{trunc}} \nonumber
\end{eqnarray}
So now we have
\begin{eqnarray}
s_i  & = & \tintegral \, K \, (\tage-t) \, e^{(t-\tage) /\tau}\, \frac{\log t_{i+1} - \log t}{\dlt}  \nonumber \\
      & = & \frac{K \, e^{-\tage/\tau}}{\dlt} \tintegral \, \left(\tage \, \log t_{i+1} \, e^{t/\tau}  - \tage \, \log t \, e^{t/\tau} - t \, \log t_{i+1} \, e^{t/\tau} + t \, \log t \, e^{t/\tau}\right) \quad .\nonumber
\end{eqnarray}
The first and third terms are fairly straightforward.  The second term is basically the integral we did for the $\tau$-model.  The last term can be solved by integrating by parts (with $u=t\,\log t$, $dv=e^{t/\tau}dt$ and hence $du=(\log e + \log t)\, dt$, $v=\tau e^{t/\tau}$.  Substituting and collecting terms gives, finally, 
\begin{eqnarray}
s_i  & = & \frac{K \, \tau \,  e^{-\tage/\tau}}{\dlt} \left[\left.\left(t\,\log \frac{t}{t_{i+1}} + \tage\,\log t_{i+1}  + \tau\,\log \frac{t_{i+1}}{e}\right) e^{t/\tau}\tinterval - (\tage/\tau+1)\, H_i\right] \quad .
\end{eqnarray}

For interpolation in linear time, we have
\begin{eqnarray}
g_i  & = & \tintegral \, K \, (\tage-t) \, e^{(t-\tage) /\tau}\, \frac{t_{i+1} - t}{\dt}  \nonumber \\
      & = & \frac{K \, e^{-\tage/\tau}}{\dt} \tintegral \, \left(\tage \, t_{i+1} \, e^{t/\tau}  - \tage \, t \, e^{t/\tau} - t \,  t_{i+1} \, e^{t/\tau} + t^2 \, e^{t/\tau}\right) \nonumber \\
     & = & \frac{K \, \tau \, e^{-\tage/\tau}}{\dt} \left. e^{t/\tau} \, \left[ \tage \, t_{i+1} - (\tage + t_{i+1})\, (t-\tau) + t^2 - 2\, t\, \tau + 2\, \tau^2\right] \tinterval \quad .
\end{eqnarray}

\subsection{Simha}
This is the SFH from \citet{simha14}.  It is defined by a delayed-$\tau$ SFH until some time $\sftrunc$, after which the SFH is linearly increasing or decreasing with some slope $m_{\mathrm{SF}}$.  The SFR is continuous at $\sftrunc$, but the derivative is not.  For this reason, we will treat the SFH in two separate pieces.  The trick is to get the proper normalization at $\sftrunc$ and, for linearly decreasing SF, find the time where the SF reaches zero. 


\subsection{Bursts}

\end{document}